Relating two Abstractions for Formal Array Languages

### 2014-04-11 paper.tex
Removed some comments in the code to save space. More text about shapes

\begin{figure}
\begin{verbatim}
-- Mapping a function over an array
map :: (a -> b) -> Pull a -> Pull b
map f (Pull ixf l) = Pull (f . ixf) l

-- Apply a binary operation itemwise on two arrays
zipWith :: (a -> b -> c)
        -> Pull a -> Pull b -> Pull c
zipWith f (Pull ixf1 l1) (Pull ixf2 l2)
  = Pull (\i -> f (ixf1 i) (ixf2 i)) (min l1 l2)

-- Takes the n first elements of an array
take :: Int -> Pull a -> Pull a
take n (Pull ixf l) = Pull ixf (min n l)

-- Drops the n first elements of an array
drop :: Int -> Pull a -> Pull a
drop n (Pull ixf l)
  = Pull (\i -> ixf (i + n)) (max (l - n) 0)

-- Rotate an array k steps to the left
rotate :: Int -> Pull a -> Pull a
rotate k (Pull ixf l)
  = Pull (\i -> ixf ((i + k) `mod` l)) l

@@

Fusion is a powerful optimization which allows programming on a high
level by composing functions without paying a performance penalty.

@@

@@

Below we show how higher dimensional Pull arrays can be defined using
the \verb!Shape! type. We also show some functions adapted for the new
representation.
\begin{verbatim}
data Pull sh a = Pull (Shape sh -> a) (Shape sh)

@@

zipWith :: (a -> b -> b) -> 
           Pull sh a -> Pull sh b -> Pull sh c
zipWith f (Pull ixf1 sh1) (Pull ixf2 sh2)
  = Pull (\sh -> f (ixf1 sh) (ixf2 sh2))
\end{verbatim}

>>>>

\begin{figure}
\begin{verbatim}
map :: (a -> b) -> Pull a -> Pull b
map f (Pull ixf l) = Pull (f . ixf) l

zipWith :: (a -> b -> c)
        -> Pull a -> Pull b -> Pull c
zipWith f (Pull ixf1 l1) (Pull ixf2 l2)
  = Pull (\i -> f (ixf1 i) (ixf2 i)) (min l1 l2)

take :: Int -> Pull a -> Pull a
take n (Pull ixf l) = Pull ixf (min n l)

drop :: Int -> Pull a -> Pull a
drop n (Pull ixf l)
  = Pull (\i -> ixf (i + n)) (max (l - n) 0)

rotate :: Int -> Pull a -> Pull a
rotate k (Pull ixf l)
  = Pull (\i -> ixf ((i + k) `mod` l)) l

reverse :: Pull a -> Pull a
reverse (Pull ixf l) = Pull (\i -> ixf (l - i - 1) l

@@

@@

\subsection{Implementing Fusion}

@@

Below we show how higher dimensional Pull arrays can be defined using
the \verb!Shape! type. We also show some functions adapted for the new
representation. Note the shape polymorphism in the definitions of
\verb!map! and \verb!zipWith!. They can work with arrays of any
shapes, as long as they are the same shape.  The function
\verb!reverse! is particularly noteworthy because the type reflects
that the input and output arrays need to have at least one dimension.

\begin{verbatim}
data Pull sh a = Pull (Shape sh -> a) (Shape sh)

@@

zipWith :: (a -> b -> b) -> 
           Pull sh a -> Pull sh b -> Pull sh c
zipWith f (Pull ixf1 sh1) (Pull ixf2 sh2)
  = Pull (\sh -> f (ixf1 sh) (ixf2 sh2))

reverse :: Pull (sh :. Data Length) a ->
           Pull (sh :. Data Length) a
reverse (Pull ixf sh@(_ :. l)) =
 Pull (\(sh' :. i) -> ixf (sh' :. (l - i - 1))) sh
\end{verbatim}
